Analysing Melbourne CBD pedestrian counts data with a focus on imputation methods

### 2017-10-23 thesis_draft/01-chap1.Rmd
pdf with png

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE,
                      message = FALSE, warning = FALSE,
                      fig.pos = 'h')
library(tidyverse)
library(lubridate)
library(foreach)
library(doSNOW)
library(rwalkr)
library(pander)
library(RColorBrewer)
library(gridExtra)
library(viridis)
library(ggmap)
library(broom)
set.seed(12345)
```

@@

The main issue which is encountered with traffic data in general is the presence of missing values. For the Melbourne CBD pedestrian data, data is missing as a result of the

@@

# Motivation
Modelling pedestrian activity in the Melbourne CBD will provide insights to the social and consumer behaviour of the people within the city. In particular, prediction of pedestrian traffic will provide information about the way the city operates. Examples of some uses of such information include infrastructure planning, or security planning (which has become quite important in recent times) within the government sector. Private uses of such information include marketing, such as finding optimal locations to launch advertising campaigns, resource management for businesses to improve staffing based on pedestrian traffic, and investment planning to analyse business plan feasibility.

@@

The first method used to impute missing values is to fit generalised linear models (GLMs) at each sensor location. Estimating a GLM with a Quasi-Poisson error distribution, where specify the model as:
$$E \left[\text{HourlyCounts} \vert \text{Sensor}, \text{HDay}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{HDay}, \text{Time}}) $$
$$\text{where } \quad \mu_{\text{Sensor}, \text{HDay}, \text{Time}} \sim \text{Time}\times \text{HDay} \tag{1}$$
`Time` is the time of the day, while `HDay` is the day of the week, with an additional factor level for public holidays. 
Both these variables are categorical/factors. Time of the day  cannot be treated as an integer or numeric because it does not have a linear relationship with the pedestrian counts.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE,
                      message = FALSE, warning = FALSE,
                      fig.pos = 'h', dev = 'png')
library(tidyverse)
library(lubridate)
library(foreach)
library(doSNOW)
library(rwalkr)
library(pander)
library(RColorBrewer)
library(gridExtra)
library(viridis)
library(ggmap)
library(broom)
set.seed(12345)
```

@@

The main issue which is encountered with traffic data in general is the presence of missing values. For the Melbourne CBD pedestrian data, data is missing as a result of sensors being installed over time, as well as technical issues. As a consequence, data imputation, the estimation of missing data, is required.

@@

## Motivation

Modelling pedestrian activity in the Melbourne CBD will provide insights to the social and consumer behaviour of the people within the city. In particular, prediction of pedestrian traffic will provide information about the way the city operates. Examples of some uses of such information include infrastructure planning, or security planning (which has become quite important in recent times) within the government sector. Private uses of such information include marketing, such as finding optimal locations to launch advertising campaigns, resource management for businesses to improve staffing based on pedestrian traffic, and investment planning to analyse business plan feasibility.

@@

The first method used to impute missing values is to fit generalised linear models (GLMs) at each sensor location. Estimating a GLM with a quasipoisson error distribution, where specify the model as:
$$E \left[\text{HourlyCounts} \vert \text{Sensor}, \text{HDay}, \text{Time} \right] = \exp (\mu_{\text{Sensor}, \text{HDay}, \text{Time}}) $$
$$\text{where } \quad \mu_{\text{Sensor}, \text{HDay}, \text{Time}} \sim \text{Time}\times \text{HDay} \tag{1}$$
`Time` is the time of the day, while `HDay` is the day of the week, with an additional factor level for public holidays. 
Both these variables are categorical/factors. Time of the day  cannot be treated as an integer or numeric because it does not have a linear relationship with the pedestrian counts.
